\documentclass[12pt]{article}
\usepackage{xcolor}
\usepackage[framemethod=TikZ]{mdframed}
\usepackage{geometry}   

\definecolor{githubgray}{RGB}{106, 115, 125}
\definecolor{githubblue}{RGB}{36, 41, 46}
\definecolor{githubquote}{RGB}{248, 249, 250}
\definecolor{bordercolor}{RGB}{230, 232, 234}

\geometry{margin=1in}

\mdfdefinestyle{rounded}{
    roundcorner=6pt,
    linewidth=1pt,
    linecolor=bordercolor,
    backgroundcolor=githubquote,
    innertopmargin=12pt,
    innerbottommargin=12pt,
    innerleftmargin=15pt,
    innerrightmargin=15pt
}

\begin{document}

\begin{center}
    {\huge\textbf{Assignment 1}}
    \vspace{1em}
    
    {\large Computer Ethics - CIS 381}
    \vspace{2em}
    
    \begin{tabular}{c}
    Yazeed AlKhalaf - 202211123 \\
    Khaled AlAnbar - 202211365
    \end{tabular}
    \vspace{2em}
\end{center}

\section{Scenario}

Students at Al Yamamah University sit in the faculty owned meeting room. This room's rules requires a faculty member to be present when a student is there. Some student sit there to study without a faculty member presence. Most of the sutdents think that this is acceptable because most of the times the room is free and they only use it when it is free. On the other hand, the rules are rules, and they say otherwise.

\subsection{Deontological View}

\begin{mdframed}[style=rounded]
An ethical philoshophy in which an action is considered morally good because of some characterusut of the action itself, not because the prodict of the action is good.
\end{mdframed}

In this scenario based on the deonotological view, students are breaking the rules of the university. Even though they are benfitting by studying when it is empty, is doesn't change the fact that they did something wrong and it is not allowed.

\subsection{Utilitarianism View}

\begin{mdframed}[style=rounded]
    An ethical philosophy in which the happiness of the greatest number of people in the society is considered the greatest good.
\end{mdframed}

In this scenario based on the utilitarianism view, it is okay to use the room since most if not all of the students are beneftiitng from the room without hurting anyone since they use it when it is free.

\end{document}